Knowledge Representation for Service Robots: A Survey

### 2014-04-16 krs-survey.tex
Fixed ORO specifics and proof-read discussion

Section~\ref{sect|features} introduces a typology of requirements and features
of knowledge representation systems for robotics, that we use as foundations to
compare and evaluate the systems we examine in this survey.

Section~\ref{sect|surveyed-systems} then presents each of the systems we have
examined, and underlines their specificities.

We finally summarize in section~\ref{sect|conclusion} the current approaches in
order to identify several research directions that are not sufficiently
addressed by the current state of the art.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Towards the next generation of Knowledge Representation Systems for Robotics}
\label{sect|conclusion}

This section tries to summary the various approaches we surveyed in the
previous sections and to draw some research perspectives. This section does not
exactly follow the list of compared features presented in
section~\ref{sect|features}. We have organized it along four axis: \emph{what
can be represented?}, \emph{How knowledge is created and grounded?}, \emph{What
can be done with the knowledge?} and \emph{How to use knowledge in the whole,
larger robot architecture?}.

The table~\ref{table|contribution-by-systems} sketches the landscape of current
research. The main fields of \emph{contribution} for the systems we have
surveyed are sorted along the different axis, dimensions of knowledge
representation that were identified in previous section.

@@

\caption{Main domains of contribution of current KRS. Italics mean that the
feature is implemented as an external module. Main references are given in the
table header. When relevant, feature-specific publications have been provided.
An empty cell means that either the system has no specific focus on this domain
or we could not find relevant literature.}

@@

To comment this table, we propose select nine topic based on some of the
challenges that McCarthy and Roy have identified (there are listed at
section~\ref{sect|evaluation-literature}) on the road to natural interaction
and, to slightly paraphrase McCarthy, ``human-level robots''.

McCarthy underlines the necessity of relying on mathematical logic, as the most
fruitful formalism for machine intelligence. This is largely the case. Almost
all the systems we have surveyed rely on logical formalisms, in several cases
as a mix of declarative languages (often based on Decription Logics) and
logical programming languages (Prolog, of course, but not only).

The two exceptions (the CAST knowledge model and GSM) are however also
interesting: CAST proposes a pervasive model of knowledge that is seducing from
the grounding perspective (but likely suboptimal for deliberative tasks), and
GSM encodes knowledge in an amodal (continuous, geometric) model while
preserving features that are usually specific of symbolic models (like theory
of mind or categorical knowledge).

The interleaving of discrete symbolic models with continuous geometric models
remains a major challenge, and techniques vary a lot: besides the CAST proxies
and the GSM amodal model, most systems directly extract physical attributes of
the environment to insert them in the symbolic model ({\sc Tapas}, Ke Jia,
OUR-K, PEIS), thus skipping altogether an intermediate geometric model. {\sc KnowRob}
adopt a top-down approach where local geometric models are set up on-demand to
compute symbolic properties like relative locations.

In our approach, ORO has tight (but unidirectional) links with SPARK, the
component in charge of geometric modeling and reasoning. Similarly to the GSM
approach, SPARK is an amodal model of the environment we can manipulate (3D
motion planning, for instance takes place in MHP, the twin brother of SPARK).

The most notable attempt at modeling uncertainty down to the knowledge
representation is {\sc ProgCog}, an extension of {\sc KnowRob}. The theoretical
and practical performances (including decidability) remain however difficult to
overcome.

The challenge of representing and reasoning under uncertainty has not been
tackled in ORO server, at least not \emph{within} the current ORO server
knowledge model. Two reasons explain this absence: the decisional architecture
developed at LAAS for the robots has not consistent approach to uncertainty
representation and management, neither at the \emph{symbol production} level
(\ie SPARK or {\sc Dialogs}) or at the symbol consumption level (execution
control). Thus, no strong incentive pushed us in this direction. Then, the
current tools available from the Semantic Web domain to manipulate and reason
about knowledge do not provide mature support for uncertainty management. Some
effort do exist (like the {\sc Pronto} reasoner~\cite{Klinov2008}), but this
does not appear to be a major focus in the currently available tools.

@@

Non-monotonicity can also be partially dealt with with appropriate time
representation and reasoning (the reasoner then only takes into account
statements that are set as valid for a given moment).

@@

Several approaches for building contextualized knowledge have been presented in
this article: symbolic environment interpretation, perspective taking and
independent mental state for each agent, grounded natural language resolution,
self-awareness of its own activity. Much remains to be done for a robot to
actually identify its current context as well as contexts that may be referred
to.

We will further develop this important topic in the conclusion of the paper.

@@

Integration of these techniques with memory mechanisms (forgetting and
reinforcement learning) remains to explore.

Temporal modeling is currently absent of ORO, and symbolic and geometric
planning are accomplished outside of the main knowledge representation
layer. As a matter of fact, knowledge is mostly atemporal in the ORO server: the
set of triples stored in the server represents the beliefs of the robot
\emph{at the current time}. The robot represents knowledge about the \emph{here
and now}. Dealing with time reasoning is delegated to dedicated tools.

However, ORO does provide a simple implementation of a memory mechanism that
relies on time: when a statement is attached to a memory profile, the statement
is reified and the date of creation is stored. This allows to remove the
statement after a period of time that depends on the memory profile.

Statement reification to store timestamps of creation for all statements is
technically very easy to do in ORO server, but has not been actually enabled
because the performance hit (each reified statement produces four triples
instead of one) was not justified by any current use case of ORO server.

@@

A side note about how thematic roles and action models are initialized in ORO:
The current implementation relies on a small, predefined set of action verbs
that can be recognised from natural language
(section~\ref{processing_of_actions}).  This constraint does not come from the
resolution algorithm itself, but rather from the difficulty to automatically
extract the thematic roles associated to a verb. An interesting yet easy to
implement extension would consist in binding the dialogue processor with the
symbolic task planner. The task planner could dynamically provide the list of
actions that the robot can process, \ie actions for which the robot can produce
a plan. In certain case, it could also validate the understanding of a desire
by attempting to plan it.

Also, we could exploit on-line resources like
\textsc{VerbNet}~\cite{Kipper2008}, which provides a large machine-processable
lexicon of English verbs along with their thematic roles.

@@

Symbol grounding remains a focus for most of the developers of cognitive
architectures for robots, and all the systems we have surveyed implement
grounding strategies (...it was an inclusion criteria). Most of the systems
adapt their grounding strategies to the sensing modality, so that grounding can
not be considered as a single, well delimited process.

@@

We note in addition that the development of what we have called ``synthetic
sensors'' is simplifying the grounding task. The most obvious example is the
human tracking system provided by the low-cost Kinect device, and now used
pervasively in the robotics lab. Not only this system segments and computes the
pose of humans that enter its field of view, but it also tracks them (including
in case they are partially or completely occluded). This kind of high level
sensing device does not completely remove the need of grounding one symbolic
instance of a human with the physical human, but the task is ways simpler than
it used to be.

Many systems also tackle the difficult question of natural language processing.
The CAST middleware, in particular, has been used in the European CoSy and CogX
projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
the huge interest for the grounding of verbal interaction within the cognitive
robotics community, and we are likely to reach important milestones in this
field in the coming years.

@@

Our knowledge representation systems have a pragmatic and potentially very
powerful approach to common-sense: reuse knowledge stored on the Web. While not
so many of the systems we have surveyed directly tackle this question ({\sc
KnowRob} is the only one explicitly working on common-sense reasoning, through
physics-based reasoning, integration with Web databases or parsing of
semi-structured Web documents), several rely however on Web standards (OWL,
OpenCyc) to represent their knowledge. With the development of initiatives like
OpenMind that encode with the same standards common-sense facts and rules,
there is a strong potential for our robot to gain common-sense knowledge from
well-structured online resources in the coming years.

@@

Table~\ref{table|knowledge-sources} lists for the systems we have surveyed the
current sources of common-sense knowledge.

@@

Learning has not been addressed in this work. This is a complex, still emerging
field, whose exact meaning and perimeter varies a lot. It was difficult to
synthesise in the KRS taxonomy, and we decided to omit it. We did not either
really tackled this question in ORO, except in term of knowledge structure
alteration based on verbal interaction (this is presented in one case-study, in
the next section).

Learning at the level of the knowledge representation system has been explored
in conjunction with visual perception systems (mentioned for instance, in the
CAST project, in~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its pool of
facts from informations automatically extracted from the Web~\cite{Nyga2009},
can also be considered as a \emph{learning} system (it autonomously acquire
knowledge). We lack however a formal study of learning strategies and
techniques at the knowledge representation level.

@@

Finally, Roy mentions \emph{the ability to take into account the human
perspective}: this cognitive ability, that relates to the representation of
different modalities, is present in ORO and GSM. ORO explicits the perspectives
in different symbolic mental states while GSM recursively stores models for
each agent.

We have already explained how perspective awareness enables advanced cognitive
capabilities like a theory of mind, and this is probably the main contribution
of this paper regarding techniques for knowledge representation.

% \subsection{Justification and resolution of cognitive conflicts}
% 
% Logical inconsistencies in one ORO's model (or, from a cognitive point of view,
% cognitive conflicts) could be better dealt with: while (partially) provided by
% the Pellet reasoner, and output to the developers for debugging purposes, the
% logical justification of the inconsistency is not really exploited.

% \subsection{Performances analysis}
% 
% \subsubsection{Relaxation time}
% 
% The relaxation time (as defined at section~\ref{sect|performances}) of ORO for
% the average workloads we have encountered during the experiments is in the
% range of 100ms.  The temporal resolution that follows for the symbolic layer,
% while acceptable in most cases, is not satisfactory. In particular, multi-modal
% interaction requires fine grained temporal alignement between
% modalities~\cite{Li2012} which are not possible with the current temporal
% performances of ORO.
% 
% \subsubsection{Cognitive evaluation}
% 
% The performance evaluation of current tools is still mostly system- and
% scenario-specific (``did my system do the job for my task?''), and amongst the
% surveyed KRS, only few systems proposes an evaluation of their performances
% that could be compared to other system.
% 
% 
% \fxwarning{Mention reasoning and expressiveness, cf section oro-server}
%

\subsection{Towards a new design}

>>>>

Section~\ref{sect|krsoverview} gives a first broad overview of knowledge
representation in robotics. It introduces briefly a range of different
approaches that do not focus on a specific field of robotics, and instead
provide insights on the wealth of knowledge representation techniques available
to researchers.

Sections~\ref{sect|features} and~\ref{sect|surveyed-systems} then focus on eight
systems deployed on \emph{service robots}: Section~\ref{sect|features}
introduces a extensive typology of requirements and features of knowledge
representation systems, that we use as foundation to compare and evaluate the
systems we examine, and section~\ref{sect|surveyed-systems} then presents each
of the eight systems, and underlines their specificities.

We finally summarize in section~\ref{sect|conclusion} the current approaches and
we attempt to identify several research directions that are not sufficiently
addressed by the current state of the art.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Knowledge Representation in Robotics: an Overview}
\label{sect|krsoverview}

TDB

\section{Evaluation Criteria: A Typology of Knowledge Representation Requirements for Robotics}
\label{sect|features}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis: How Current Systems Tackle the Knowledge Representation
Challenge?}
\label{sect|conclusion}

This section summarizes the different approaches to knowledge representation and
reasoning that are surveyed in the previous section, and attempt to identify
research perspectives.

First, the Table~\ref{table|contribution-by-systems} presents, system by system
and feature by feature, a landscape of current research, underlining the main
design choices and scientific \emph{contribution} of each of the KRS that we
have examined.

Then, we go over again the nine ``challenges'' proposed by McCarthy and Roy (see
section~\ref{sect|evaluation-literature}) to evaluate how far we are on the road
to natural interaction and, to slightly paraphrase McCarthy, ``human-level
robots''.

McCarthy underlines the necessity of relying on mathematical logic, as the most
fruitful formalism for machine intelligence. This is largely the case. Almost
all the systems we have surveyed rely on logical formalisms, in several cases
as a mix of declarative languages (often based on Decription Logics) and
logical programming languages (Prolog, expectedly, but not only).

The two exceptions (the CAST knowledge model and GSM) are however 
interesting: CAST proposes a pervasive model of knowledge that is convincing from
the grounding perspective (but likely suboptimal for deliberative tasks), and
GSM encodes knowledge in an amodal (continuous, geometric) model while
preserving features that are usually specific of symbolic models (like theory
of mind or categorical knowledge).

The interleaving of discrete symbolic models with the continuous physics of the
world remains a major challenge, and techniques vary a lot: besides the CAST
proxies and the GSM amodal model, most systems directly extract physical
attributes of the environment to insert them in the symbolic model ({\sc Tapas},
Ke Jia, OUR-K, PEIS), thus skipping altogether an intermediate geometric model.
ORO has tight (but unidirectional) links with an intermediate geometric
representation (called {\sc Spark}), and {\sc KnowRob} adopt a top-down approach
where local geometric models are set up on-demand to compute symbolic properties
like relative locations.

The most notable attempt at modeling uncertainty down to the knowledge
representation is {\sc ProgCog}, an extension of {\sc KnowRob}. Theoretical
and practical performances issues (including decidability) remain however
difficult to overcome.

It can be also noted that some effort do exist to provided probabilistic
reasoning to Description Logics based systems (like the {\sc Pronto}
reasoner~\cite{Klinov2008}), but this does not appear to be a major focus in the
currently available tools.

@@

\caption{Design choices and main domains of contribution of current KRS (+
indicates \emph{moderate contribution}, +++ indicates \emph{major
contribution}). Italics mean that the feature is implemented as an external
module. Main references are given in the table header. When relevant,
feature-specific publications have been provided.  An empty cell means that
either the system has no specific focus on this domain or we could not find
relevant literature.}

@@

@@

Non-monotonicity can also be partially dealt with with appropriate time
representation and reasoning (the reasoner then only takes into account
statements that are set as valid for a given moment) or with dedicated reasoners
for tasks requiring non-monotonic reasoning like symbolic task planning
\fxwarning{mention ORO + HATP}.

@@

Several approaches for building contextualized knowledge have been however
hinted in section~\ref{sect|features}: symbolic environment interpretation,
perspective taking and independent mental state for each agent, grounded natural
language resolution, self-awareness of its own activity. Much remains to be done
for a robot to actually identify its current context as well as contexts that
may be referred to.

@@

Integration of these techniques with memory mechanisms (forgetting and
reinforcement learning) remains mostly to be explored.

@@

@@

Symbol grounding remains a focus for most of the developers of cognitive
architectures for robots, and all the systems we have surveyed implement
grounding strategies. Most of the systems
adapt their grounding strategies to the sensing modality, so that grounding can
not be considered as a single, well delimited process.

@@

We note in addition that the development of high-level sensors is simplifying
the grounding task. The most obvious example is the human tracking system
provided with the low-cost Kinect family of devices, and now used
pervasively in the robotics lab. Not only this system segments and computes the
pose of humans that enter its field of view, but it also tracks them (including
in case they are partially or completely occluded). This kind of high level
sensing device does not completely remove the need of grounding one symbolic
instance of a human with the physical human, but the task is much simpler than
it used to be.

Many systems also tackle the difficult question of natural language processing.
The CAST middleware, in particular, has been used in the European CoSy and CogX
projects as knowledge base by linguists~\cite{Kruijff2010a}. This demonstrates
the interest for the grounding of verbal interaction within the cognitive
robotics community, and we are likely to reach important milestones in this
field in the coming years.

@@

Knowledge representation systems have usually a pragmatic approach to
common-sense: they reuse knowledge bases stored on the Web. While {\sc KnowRob}
is the only one explicitly working on common-sense reasoning, through
physics-based reasoning, integration with Web databases or parsing of
semi-structured Web documents), several other KRS rely on Web standards (OWL,
OpenCyc) to represent their knowledge (Table~\ref{table|knowledge-sources} lists
the sources of common-sense knowledge found in the surveyed systems). With the
development of initiatives like OpenMind that encode with the same standards
common-sense facts and rules, there is a strong potential for our robot to gain
common-sense knowledge from well-structured online resources in the coming
years.

@@

@@

Learning at the level of the knowledge representation system has been explored
in conjunction with visual perception systems (mentioned for instance for the
CAST project~\cite{jacobsson2008crossmodal}). {\sc KnowRob}, by filling its pool of
facts from informations automatically extracted from the Web~\cite{Nyga2009},
can also be considered as a \emph{learning} system (it autonomously acquire
knowledge). We lack however a formal study of learning strategies and
techniques at the knowledge representation level.

@@

Finally, Roy mentions \emph{the ability to take into account the human
perspective}: this cognitive ability, that relates to the representation of
different modalities, is explicitly advertised in ORO and GSM only: ORO
explicits the perspectives in different symbolic mental states while GSM
recursively stores models for each agent.  We think however that several of the
other systems could easily gain support for alternative belief models, leading
the way to becoming perspective-aware KRS.

\subsection{Conclusion: the Future Directions to Explore}

TDB
